Tulai

### 2023-09-29 paper/paper.qmd
stash

```{r}
# load libraries
library(tidyverse)
library(here)
library(ggpmisc)
library(ggrepel)
library(ggbeeswarm)
library(cowplot)
library(knitr)
library(ggcorrplot)
library(ggpubr)
library(ggbiplot)
```

@@

# chi-square test for different frequencies of tool type groups 
# by area
  summary_table_area_filtered_for_chi_square <- 
  summary_table_area_filtered %>% 
  pivot_longer(-area) %>% 
  pivot_wider(names_from = "area") %>% 
  select(-name)

chisq <- 
  chisq.test(summary_table_area_filtered_for_chi_square)

rownames(chisq$residuals) <- 
   summary_table_area_filtered %>% 
  pivot_longer(-area) %>% 
  pivot_wider(names_from = "area") %>% 
  pull(name)  %>% 
  str_remove("\\(n\\)")

ggcorrplot(chisq$residuals,
           method = 'circle') + 
  ggplot2::scale_fill_gradient2(
    low = 'red',
    high = 'royalblue',
    mid = 'white',
    midpoint = 0,
    limit = c(-7, 11),
    space = "Lab",
    name = "Residual"
  ) + 
  coord_flip()

@@

A Chi-square test on the frequency of major lithic types by excavation area reveals a significant difference between the areas (`r apastats::describe.chi(summary_table_area_filtered_for_chi_square)`). Figure XXX shows that there are fewer utlized pieces and flakes in d1 than expected, and more flakes and fewer cores and blades in tp1 than expected assuming a null hypothesis of no difference between the areas.

@@

kable(
  summary_table_tool_area,
  col.names = c("Tools", "D1", "TA", "TP1", "Total"),
  caption = "Tool Typology in the three main areas")
```

# chi-square test for different frequencies of tool type groups 
# by area
  summary_table_tool_area_for_chi_square <- 
  summary_table_tool_area %>% 
  filter(!tooltypegroup == "Total") %>% 
  select(d1, ta, tp1)

chisq <- 
  chisq.test(summary_table_tool_area_for_chi_square)

rownames(chisq$residuals) <- 
  summary_table_tool_area %>% 
  filter(!tooltypegroup == "Total") %>% 
  pull(tooltypegroup)

ggcorrplot(chisq$residuals,
           method = 'circle') + 
  ggplot2::scale_fill_gradient2(
    low = 'red',
    high = 'royalblue',
    mid = 'white',
    midpoint = 0,
    limit = c(-7, 11),
    space = "Lab",
    name = "Residual"
  ) + 
  coord_flip()

@@

A Chi-square test on the frequency of tool type groups by excavation area reveals a significant difference between the areas (`r apastats::describe.chi(summary_table_tool_area_for_chi_square)`). Figure XXX shows that there are more geomeric pieces and burins in d1 than expected, and more backed and retouched pieces in tp1 than expected assuming a null hypothesis of no difference between the areas.

@@

# write the percentage table to a CSV file
write_csv(summary_table_tool_area_percentage,
          here("data/derived-data/tbl-002-summary_table_tool_area_percentage.csv"))

kable(
  summary_table_tool_area_percentage,
  col.names = c("Tools", "D1 (%)", "TA (%)", "TP1 (%)", "Total (%)"),
  caption = "Tool Typology in the three main areas (Percentage)"
)

@@

# Generate the base biplot
biplot_base <- fviz_pca_biplot(res.pca,
                               axes = c(1, 2),
                               labelsize = 2,
                               label = "var",
                               addEllipses = TRUE,
                               col.ind = excavation_area,
                               title = ""
                              )

# Modify row names of coordinates
rownames(res.pca$ind$coord) <- sum_depth_sheet_subset$label2

pca_label_text <- 
res.pca$ind$coord %>% 
  as_tibble() %>% 
  mutate(labels = sum_depth_sheet_subset$label2,
         area =   excavation_area)

# Add geom_text_repel() to the base plot
biplot_modified <- biplot_base +
                   geom_text_repel(data = pca_label_text,
                                   aes(x = Dim.1, 
                                       y = Dim.2, 
                                       label = labels,
                                       colour = area),
                                   segment.color = "grey80",
                                   size = 2,
                                   force = 10
                                  ) +
  labs(colour = "",
       shape = "",
       fill = "") +
  theme(legend.position = c(0.9, 0.2))

@@

ggsave(here("figures/fig-pca-biplot.jpg"),
       h = 6, 
       w = 8)

>>>>

```{r}
# load libraries
library(tidyverse)
library(here)
library(ggpmisc)
library(ggrepel)
library(ggbeeswarm)
library(cowplot)
library(knitr)
```

@@

@@

@@

kable(
  summary_table_tool_area,
  col.names = c("Tools", "D1", "TA", "TP1", "Total"),
  caption = "Tool Typology in the three main areas")

@@

@@

# write the percentage table to a CSV file

@@

library(ggpubr)
library(ggbiplot)

pca_data.pca <- prcomp(pca_data, scale. = TRUE)

biplot_modified <- 
ggbiplot(
  pca_data.pca,
  obs.scale = 1,
  var.scale = 1,
  groups = excavation_area,
  ellipse = FALSE,
  circle = TRUE
) +
  stat_chull(aes(color = excavation_area,
                 fill = excavation_area),
             alpha = 0.1,
             geom = "polygon") +
  scale_colour_brewer(palette = "Set1",
                      name = '',
                      guide = 'none') +
  scale_fill_brewer(palette = "Set1",
                    name = '')  +
  theme_minimal()

@@

ggsave(plot = biplot_modified,
       filename = here::here("figures/fig-pca-biplot.jpg"),
       h = 6, 
       w = 8)
